---
  title: "Genome Annotation Project"
subtitle: "Figure 8: undetected CDS in UTRs"
date: "April 10, 2024"
author: "Adam Freedman"
output: 
  html_document: default
---
  
## Set up
```{r,,message=FALSE, warning=FALSE}
currentdir=getwd()
setwd(currentdir)
library(tidyverse)
library(scales)
``` 

## Data
data <- read_csv("methodsintegration.csv")

## Figure 9A
```{r}
methodlabels<-c(expression(Stringtie[STAR]),"Stringtie+TOGA",expression(paste("Stringtie+TOGA+",BRAKER[protein])))
stie1st_biplot <- data %>% filter(method %in% c("stiestar","stietoga","stietogabprot")) %>%
      mutate(method=factor(method,levels=c("stiestar","stietoga","stietogabprot")))%>%
      ggplot(aes(x=busco_score,y=align_rate,color=method)) + 
      geom_point(alpha=0.5,size=1.5) + 
      ylim(0.75,0.78) +
      scale_color_manual(labels=methodlabels,values=c("magenta","dodgerblue","blue")) +
      theme(text=element_text(size=8),legend.text.align = 0,legend.spacing.y = unit(0.15, "cm"),legend.key.size= unit(0.5, "cm")) +
      guides(color = guide_legend(title = "Method")) +
      xlab("BUSCO score") +
      ylab(expression(paste(RNA[seq], " alignment rate"))) 
print(stie1st_biplot)
ggsave(stie1st_biplot,file="Figure_9A.pdf",width=5.05,height=1.5)
```

## Figure 9B
```{r}
methodlabels<-c("TOGA","TOGA+Stringtie",expression(paste("TOGA+Stringtie+",BRAKER[protein])))
toga1st_biplot <- data %>% filter(method %in% c("toga","togastie","togastiebprot")) %>%
      mutate(method=factor(method,levels=c("toga","togastie","togastiebprot")))%>%
      ggplot(aes(x=busco_score,y=align_rate,color=method)) + 
      geom_point(alpha=0.5,size=1.5) + 
      scale_color_manual(labels=methodlabels,values=c("black","dodgerblue","blue")) +
      theme(text=element_text(size=8),legend.text.align = 0,legend.spacing.y = unit(0.15, "cm"),legend.key.size= unit(0.5, "cm")) +
      guides(color = guide_legend(title = "Method")) +
      xlab("BUSCO score") +
      ylab(expression(paste(RNA[seq], " alignment rate"))) 
print(toga1st_biplot)
ggsave(toga1st_biplot,file="Figure_9B.pdf",width=5,height=1.5)
```

## Figure 9C
```{r}
methodlabels<-c(expression(Stringtie[STAR]),expression(paste("Stringtie+",BRAKER[RNA])),expression(paste("Stringtie+",BRAKER[RNA],"+",BRAKER[protein])))
notoga_biplot <- data %>% filter(method %in% c("stiestar","stiebrna","stiebrnabprot"))  %>%
      mutate(method=factor(method,levels=c("stiestar","stiebrna","stiebrnabprot")))%>%
      ggplot(aes(x=busco_score,y=align_rate,color=method)) + 
      geom_point(alpha=0.5,size=1.5) + 
      scale_color_manual(labels=methodlabels,values=c("magenta","dodgerblue","blue")) +
      theme(text=element_text(size=8),legend.text.align = 0,legend.spacing.y = unit(0.15, "cm"),legend.key.size= unit(0.5, "cm")) +
      guides(color = guide_legend(title = "Method")) +
      xlab("BUSCO score") +
      ylim(0.76,0.77) +
      ylab(expression(paste(RNA[seq], " alignment rate"))) 
print(notoga_biplot)
ggsave(notoga_biplot,file="Figure_9C.pdf",width=5.375,height=1.5)
```

## Figure 9D
```{r}
methodlabels<-c("TOGA",expression(paste("TOGA+",BRAKER[protein])))
proteinonly_biplot <- data %>% filter(method %in% c("toga","togabprot")) %>%
      mutate(method=factor(method,levels=c("toga","togabprot"))) %>%
      ggplot(aes(x=busco_score,y=align_rate,color=method)) + 
      geom_point(alpha=0.5,size=1.5) + 
      scale_color_manual(labels=methodlabels,values=c("black","blue")) +
      theme(text=element_text(size=8),legend.text.align = 0,legend.spacing.y = unit(0.15, "cm"),legend.key.size= unit(0.5, "cm")) +
      guides(color = guide_legend(title = "Method")) +
      xlab("BUSCO score") +
      ylab(expression(paste(RNA[seq], " alignment rate"))) 
print(proteinonly_biplot)
ggsave(proteinonly_biplot,file="Figure_9D.pdf",width=4.75,height=1.5)
```